### 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: refactor multibase

\section{Auto-scaling -- single-provider based}
\subsection*{Description}
\begin{asparaenum}
\item[\textbf{Motivation}]The aim of this test-case is to show that introducing a multi-layered auto-scaling platform is of a great benefit in terms of cost and resource usage for cloud consumers and cloud providers respectively. What is more, this test should prove that once there are many levels on which scaling operations can be performed, there are significant reduction in costs paid by consumers.
We want to prove it by comparing our proof-of-concept product to \emph{Carina} \cite{Carina}. \emph{Carina} embraces a whole range of various scaling policies for users' environments (environment, in this context, means an application with all its software dependencies that are to be deployed on the cloud), but there is only one way in which they are executed -- by managing the number of virtual machines (i.e. horizontal scaling). Quite on the contrary, \emph{Cloud-SAP} has mechanisms that allow to scale the environment vertically in the first place and if that turns out to be not sufficient, horizontally. This test shows the influence of lack/presence of this feature on price and resource consumption.
\item[\textbf{Scenario}] The test scenario involves
  \begin{inparaenum}[a)]
    \item deployment of a sample environment on the cloud,
    \item substitute the real module responsible for collecting CPU usage date for a mock one,
    \item monitoring the scaling actions performed by each solution,
    \item evaluation of the cost the client has to pay for the service.
  \end{inparaenum}
  Deployment of a service is done in a product-specific manner (up to the point where the vm deployment request is passed to OpenNebula). Mocking the CPU usage was possible by replacing the part in \emph{InformationManager} responsible the collecting CPU data in a host for a request to a web service which generated a time-based mocked-values. Choosing the appropriate function is another issue and is discussed in the next paragraph.
\end{asparaenum}

@@

\begin{asparaenum}
  \item[\textbf{Carina}] As \emph{Carina} is capable only of horizontal scaling, in discussion of its resource usage we can confine ourselves to counting the number of deployed virtual machines in the given time intervals. In the log files we can trace all actions that were triggered during the test case. As listing \ref{lst:auto-scaling-1cp-results-carina-jobs} shows, they completely met the expectations expressed in the previous section -- there were two ''SCALEUP'' jobs which resulted in increase of the total number of virtual machines comprising the environment.  \lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-1cp-results-carina-jobs]{auto-scaling-1cp-results-carina-jobs}. Since the base configuration assumed that the environment consists of 2 virtual machines, we can say that about 10 minutes after deployment of an environment, the number of VMs increased to 3, and after another 10 minutes, to 4.
  \item[\textbf{Cloud-SAP}] Our proof-of-concept solution behaved exactly as expected. To make the discussion more clear, there is an excerpt from a log file of this test run in listing \ref{lst:auto-scaling-1cp-results-csap-logs}. \lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions, label=lst:auto-scaling-1cp-results-csap-logs]{auto-scaling-1cp-results-csap-logs}

@@

\section{Auto-scaling -- multiple-provider based}
\subsection*{Description}
\subsection*{Preconditions}
\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
  \end{center}
  \caption{Auto-scaling - multiple-provider based: environment configuration}
  \label{eval:auto-scaling-2cp-physical-nodes}
\end{figure}

\subsection*{Results}
\subsection*{Conclusion}

\section{Deployment time -- solution comparison}
\subsection*{Description}
In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center and \emph{OpenVZ} as a hypervisor in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
\emph{Carina} \cite{Carina} can be considered a perfect match of a solution for such a comparison and tests are ran against it.

@@

- TODO dodać konfigurację każdej z maszyn wirtualnych

>>>>

\newpage
\section{Auto-scaling -- single-provider based}
\subsection*{Description}
\begin{asparaenum}
\item[\textbf{Motivation}]The aim of this test-case is to show that introducing a multi-layered auto-scaling platform is of a great benefit in terms of cost and resource usage for cloud consumers and cloud providers respectively. What is more, this test should prove that once there are many levels on which scaling operations can be performed, there are significant reduction in costs paid by consumers.
We want to prove it by comparing our proof-of-concept product to \emph{Carina} \cite{Carina}. \emph{Carina} embraces a whole range of various scaling policies for users' environments (environment, in this context, means an application with all its software dependencies that are to be deployed on the cloud), but there is only one way in which they are executed -- by managing the number of virtual machines (i.e. horizontal scaling). Quite on the contrary, \emph{Cloud-SAP} has mechanisms that allow to scale the environment vertically in the first place and if that turns out to be not sufficient, horizontally. This test shows the influence of lack/presence of this feature on price and resource consumption.
\item[\textbf{Scenario}] The test scenario involves
  \begin{inparaenum}[a)]
    \item deployment of a sample environment on the cloud,
    \item substitute the real module responsible for collecting CPU usage date for a mock one,
    \item monitoring the scaling actions performed by each solution,
    \item evaluation of the cost the client has to pay for the service.
  \end{inparaenum}
  Deployment of a service is done in a product-specific manner (up to the point where the vm deployment request is passed to OpenNebula). Mocking the CPU usage was possible by replacing the part in \emph{InformationManager} responsible the collecting CPU data in a host for a request to a web service which generated a time-based mocked-values. Choosing the appropriate function is another issue and is discussed in the next paragraph.
\end{asparaenum}

@@

\begin{asparaenum}

  \item[\textbf{Cloud-SAP}] Our proof-of-concept solution behaved exactly as expected. To make the discussion more clear, there is an excerpt from a log file of this test run in listing \ref{lst:auto-scaling-1cp-results-csap-logs}. \lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions, label=lst:auto-scaling-1cp-results-csap-logs]{auto-scaling-1cp-results-csap-logs}

\item[\textbf{Carina}] As \emph{Carina} is capable only of horizontal scaling, in discussion of its resource usage we can confine ourselves to counting the number of deployed virtual machines in the given time intervals. In the log files we can trace all actions that were triggered during the test case. As listing \ref{lst:auto-scaling-1cp-results-carina-jobs} shows, they completely met the expectations expressed in the previous section -- there were two ''SCALEUP'' jobs which resulted in increase of the total number of virtual machines comprising the environment.  \lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-1cp-results-carina-jobs]{auto-scaling-1cp-results-carina-jobs}. Since the base configuration assumed that the environment consists of 2 virtual machines, we can say that about 10 minutes after deployment of an environment, the number of VMs increased to 3, and after another 10 minutes, to 4.

@@

\newpage
\section{Auto-scaling -- multiple-provider based}
\subsection*{Description}

\subsubsection{Motivation} This test case aims to prove that scaling across multiple cloud providers is vitally important while ensuring appropriate Quality-of-Service, especially in cases of increased number of service requests. Such scaling scenario, known also as cloud-bursting, leverage benefits arising from offloading application load to an external provider. In order to verify our concept we compare number of transactions per second guaranteed by \emph{Cloud-SAP} and \emph{Carina}. While our proof of concept solution features multi-layer scaling and is cloud federation aware, \emph{Carina} adopts only horizontal scaling. It is expected that test case prove benefits emerging from enriching application platform provider with an cloud federation awareness.

\subsubsection{Scenario}
Testing requires following steps to be done:
\begin{enumerate}
\item deploy exemplary service
\item observe system behaviour under load, simulated by a resource usage mock. It is vital for a resource usage to exceed a single provider capabilities
\item assess system characteristics, including number of handled transactions per second
\end{enumerate}

\subsubsection*{Load simulation}
Similarly to a previous test case, we leveraged a resource mock that allows us to precisely control monitoring information returned to an system. Besides, in order to simplify test case, we were solely focused on a CPU usage. CPU usage function has to exceed upper threshold limit at some point and stay at that level, triggering successive auto-scaling events. However, at some point single cloud provider resources will be surpassed. Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned scenario and is illustrated in figure \ref{eval:auto-scaling-2cp-cpu-usage-function}.
\begin{equation}
 f(x) = \frac{29}{1500}x^3-\frac{21}{20}x^2+\frac{533}{30}x
 \label{eq:cpu-mock-usage-2cp}
\end{equation}

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-7/cpu-usage-mock-multiple}
  \end{center}
  \caption{Auto-scaling - multiple-provider based: CPU usage function}
  \label{eval:auto-scaling-2cp-cpu-usage-function}
\end{figure}

\subsection*{Preconditions}
\subsubsection*{Environment specification}
\begin{asparaenum}
  \item[\textbf{Auto-scaling policy specifications}] Policies used in this test are based on a threshold model with the following properties:
  \begin{itemize}
   \item value below 20 triggers scaling down event
   \item value grater than 60 triggers scaling up event
  \end{itemize}
  However, considering the fact that we are entirely focused on scaling up, only the upper limit is relevant in our case. Listings \ref{lst:auto-scaling-2cp-carina-service-spec} and \ref{lst:auto-scaling-2cp-cloud-sap-service-spec} presents scaling policies for \emph{Carina} and \emph{Cloud-SAP} respectively.

  \lstinputlisting[caption=\emph{Carina} service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-carina-service-spec]{auto-scaling-2cp-carina-service-spec}
  \lstinputlisting[caption=\emph{Cloud-SAP} service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-cloud-sap-service-spec]{auto-scaling-2cp-cloud-sap-service-spec}

  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] 
  Components common to \emph{Carina} and \emph{Cloud-SAP}, namely \emph{OpenNebula} instances and computing nodes, were configured in the same fashion. Listing \ref{lst:auto-scaling-2cp-one-config} depicts key configuration elements of \emph{OpenNebula} monitoring mechanism.

  \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-one-config]{auto-scaling-2cp-one-config}

  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:auto-scaling-2cp-carina-deployment-diagram} and \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram}.

  \begin{figure}[!ht]
    \begin{center}
      \includegraphics{chapter-7/autoscale-2cp-cloud-sap-deployment-diagram}
    \end{center}
    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
    \label{eval:autoscale-2cp-cloud-sap-deployment-diagram}
  \end{figure}

  \begin{figure}[!ht]
    \begin{center}
      \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
    \end{center}
    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
    \label{eval:auto-scaling-2cp-carina-deployment-diagram}
  \end{figure}
\end{asparaenum}

\subsubsection*{Hardware/VM configuration}

Figure \ref{eval:auto-scaling-2cp-physical-nodes} depicts physical configuration of the environment. In short, setup was as follows: all \emph{Cloud-SAP} components, apart from cloud client deployed on \emph{Laptop}, were provisioned on \emph{Desktop}. \emph{Cloud Proivder 1 (CP-1)} is \emph{OpenNebula} instance known as \emph{Frontend1} which uses \emph{Node1} as a computing node and is deployed on \emph{Desktop}, while \emph{CP\-2} uses \emph{Frontend2} and \emph{Node2}. Specification of nodes is listed in table: \ref{tbl:test-deployment-time-common-hardware-configuration}. Deployment cost of a java stack was 50 and 60 using \emph{CP-1} and \emph{CP-2}, respectively.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
  \end{center}
  \caption{Auto-scaling - multiple-provider based: environment configuration}
  \label{eval:auto-scaling-2cp-physical-nodes}
\end{figure}

\subsection*{Expectations}
Taking CPU usage function into account, we can predict outcome of the test. Scaling policy is violated in 5th minute, hence, it is expected that first scaling event occur after that point. Moreover, since that moment, CPU usage remains beyond threshold implying successive scaling actions.

Due to the fact that \emph{Cloud-SAP} favours fine-grained scaling actions such as vertical scaling, it is expected that these actions occur sooner than horizontal scaling. However, at some point \emph{CP-1} capacity won't be sufficient, hence, another stack instance (one master instance, two slaves) will be deployed on \emph{CP-2}. Subsequent scaling actions will take place solely on a \emph{CP-2} 2 and will involve further vertical scaling.

Carina, on the other hand, supports solely horizontal scaling using single cloud provider. On top of that, it is assumed that succeeding slaves instances will be added to a service up to the point where Cloud Provider won't have enough resources to proceed with further scaling requests.

\subsection*{Results}

\begin{asparaenum}

  \item[\textbf{Cloud-SAP}] 
   In total, there were 13 vertical (increasing CPU limit) 2 horizontal (adding a container) scaling events. First action took place minute after first violation of scaling policy rule. Since then, slaves' CPU were successively increasing to the point were further scaling wasn't possible - 13th minute of the test. Therefore, \emph{Cloud-SAP} deployed two new slaves on \emph{CP-2}. Listing \ref{lst:auto-scaling-2cp-results-csap-logs} presents logs covering service lifecycle. \lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions on multiple providers, label=lst:auto-scaling-2cp-results-csap-logs]{auto-scaling-2cp-results-csap-logs}

  \item[\textbf{Carina}] 
  Similarly to a \emph{Cloud-SAP}, first scaling event (slave addition) was noticed in 6th minute of test. Scaling jobs were continuously invoked every 2 minutes until capacity of a \emph{CP\-1} hasn been fully exploited. Listing \ref{lst:auto-scaling-2cp-results-carina-jobs} summarises jobs performed by \emph{Carina}. \lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-2cp-results-carina-jobs]{auto-scaling-2cp-results-carina-jobs}
\end{asparaenum}

\subsubsection*{Transactions per second}
Chart \ref{eval:auto-scaling-2cp-tps-comparison} illustrates how service capabilities changed over the time. Service capability is expressed as a number of transactions per second that can be handled by a service. Please note, that this measure, while giving good insight into application potential, is hard to simulate, hence, we assumed that 1 VCPU provides enough resources to successfully serve 100 TPS. Knowing that, we roughly estimate total service capacity, which at the end of the test amounts to:
\begin{itemize}
 \item Cloud-SAP: 283 transactions per second
 \item Carina: 180 transactions per second
\end{itemize}

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-7/auto-scaling-2cp-tps-comparison}
  \end{center}
  \caption{Auto-scaling - multiple-provider based: comparison of processed transaction per second}
  \label{eval:auto-scaling-2cp-tps-comparison}
\end{figure}

\subsection*{Conclusion}
Comparing expectations with test output, one can see that expectations has been fully met. While \emph{Carina} was solely focused on horizontal scaling and bound to a single cloud provider, \emph{Cloud-SAP} first took fine-grained actions to later leverage resources of a second cloud provider.

\emph{Cloud-SAP} has been proven to be a better solution, offering greater capabilities to a service. This was possible by exploiting resources of two federated cloud providers, crucial concept behind \emph{Cloud-SAP} design.

\newpage
\section{Deployment time -- solution comparison}
\subsection*{Description}
In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center and \emph{OpenVZ} as a hypervisor in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
\emph{Carina} \cite{Carina} can be considered a perfect match of a solution for such a comparison and tests are ran against it.

@@
